Inducing Domain Specific Languages for Bayesian Program Learning

### 2018-02-06 paper.tex
symbolic regression writing and figure

\subsection{Symbolic Regression}\label{regressionSection}
We show how to use \system to infer programs containing both discrete
structure and continuous parameters. The high-level idea is to synthesize programs with unspecified-real-valued parameters, and to fit those parameters using gradient descent.
Concretely, we ask the algorithm to
solve a set of 1000 symbolic regression problems, each a polynomial of
degree 0, 1, or 2, where our observations $x$ take the form of $N$
input/output examples, which we write as $x = \left\{(i_n,o_n)
\right\}_{n\leq N}$. For example, one task is to infer a program
calculating $3x + 2$, and the observations are the input-output
examples $\left\{(-1,-1),(0,2),(1,5) \right\}$.

We initially equip our DSL learner with addition and multiplication,
along with the possibility of introducing real-valued parameters, which we write as $\mathcal{R}$.
We define the likelihood of an observation $x$ by assuming a Gaussian noise model for the input/output examples and integrate over the real-valued parameters, which we collectively write as $\vec{\mathcal{R}}$:
\begin{align*}
\log   \probability\left[\{(i_n,o_n)\}|p\right] = \log \int \mathrm{d}\vec{\mathcal{R}}\; P_{\vec{\mathcal{R}}}(\vec{\mathcal{R}})\prod_{n\leq N}\mathcal{N}(p(i_n,\vec{\mathcal{R}})|o_n)
%\text{\emph{BIC:}} &\approx 
\end{align*}
where $\mathcal{N}(\cdot |\cdot )$ is the normal density  and $P_{\vec{\mathcal{R}}}(\cdot )$ is a prior over $\vec{\mathcal{R}}$. We approximate this marginal using the BIC~\cite{Bishop:2006:PRM:1162264}:
\begin{align*}
  \log \probability[x|p]  \approx& \sum_{n\leq N}\log \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
\end{align*}
where $\vec{\mathcal{R}}^*$ is an assignment to $\vec{\mathcal{R}}$ found by performing gradient ascent on the likelihood of the observations w.r.t. $\vec{\mathcal{R}}$.

What DSL does \system learn?
The learned DSL contains templates for quadratic and linear functions,
which lets the algorithm quickly hone in on the kinds of functions that are most appropriate to this domain.
Examining the programs themselves,
one finds that the algorithm discovers representations for each of the polynomials that minimizes the number of continuous degrees of freedom:
for example, it represents the polynomial $8x^2+8x$

\begin{tabular}{ll}
  \toprule
  Primitives& $+,\times:\mathbb{R}\to \mathbb{R}\to \mathbb{R}$\\
  & $\mathcal{R}:\mathbb{R}$ (real valued parameter)\\\midrule
  Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
  \midrule
  Likelihood $\probability[x|p]$& $\propto\exp\left( - D\log N \right)\prod_{n\leq N}\mathcal{N}(p(i_n)|o_n)$
  \\\midrule
  \begin{tabular}{l}
    Subset of \\Learned DSL
    \end{tabular}&
  \begin{tabular}{ll}
    $\lambda x. \mathcal{R}\times x + \mathcal{R}$ &linear\\
    $\lambda x. \mathcal{R} + x$ &increment\\
    $\lambda x. x\times (\text{linear } x)$&$\text{quadratic}_0$\\
    $\lambda x. \text{increment }(\text{quadratic}_0\; x)$&quadratic\\
  \end{tabular}
  \\\bottomrule
  \end{tabular}

\begin{figure}
      \includegraphics[width = 9cm]{figures/polynomial.png}
\end{figure}
\subsection{String editing}\label{textSection}

>>>>

\subsection{Symbolic Regression}\label{regressionSection}
We show how to use \system to infer programs containing both discrete
structure and continuous parameters. The high-level idea is to synthesize programs with unspecified real-valued parameters, and to fit those parameters by differentiating through the program
and optimizing using gradient descent.
Concretely, we ask the algorithm to
solve a set of 400 symbolic regression problems, each a polynomial of
degree 0 to 4, where our observations $x$ are 10
input/output examples.%, which we write as $x = \left\{(i_n,o_n)
%\right\}_{n\leq N}$. %% For example, one task is to infer a program
%% calculating $3x + 2$, and the observations are the input-output
%% examples $\left\{(-1,-1),(0,2),(1,5) \right\}$.

We initially give our system only addition and multiplication,
along with the possibility of introducing real-valued parameters, which we write as $\mathcal{R}$.
We define the likelihood of an observation $x$ by assuming a Gaussian noise model for the input/output examples,
and penalize extra continuous degrees of freedom using the BIC~\cite{Bishop:2006:PRM:1162264}:
%% \begin{align*}
%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
%% \end{align*}

%% %% and integrate over the real-valued parameters, which we collectively write as $\vec{\mathcal{R}}$:
%% %% \begin{align*}
%% %% \log   \probability\left[\{(i_n,o_n)\}|p\right] = \log \int \mathrm{d}\vec{\mathcal{R}}\; P(\vec{\mathcal{R}})\prod_{n\leq N}\mathcal{N}(p(i_n,\vec{\mathcal{R}})|o_n)
%% %% %\text{\emph{BIC:}} &\approx 
%% %% \end{align*}
%% where $\mathcal{N}(\cdot |\cdot )$ is the normal density and
%% $P_{\vec{\mathcal{R}}}(\cdot )$ is a prior over
%% $\vec{\mathcal{R}}$. We approximate this marginal using the BIC~:

%% where $\vec{\mathcal{R}}^*$ is an assignment to $\vec{\mathcal{R}}$
%% found by performing gradient ascent on the likelihood of the
%% observations w.r.t. $\vec{\mathcal{R}}$.

What DSL does \system learn?
The learned DSL contains templates for polynomials of different orders,
which lets the algorithm quickly hone in on the kinds of functions that are most appropriate to this domain (Table~\ref{regressionDSL}).
Examining the programs themselves,
one finds that the algorithm discovers representations for each of the polynomials that minimizes the number of continuous degrees of freedom:
for example, it represents the linear function $-x+2$ with the program 
\texttt{(* (* (+ x }$\mathcal{R}$\texttt{) }$\mathcal{R}$\texttt{) x)}, which has two continuous degrees of freedom -- while representing quartic functions using the invented DSL primitive $f_6$ in Table~\ref{regressionDSL}.
This falls out of our Bayesian treatment of the program learning problem -- both the bias towards shorter programs (more likely under $\mathcal{D},\theta$) and the likelihood's model BIC penalty.
\begin{table}
\begin{tabular}{l}
  \toprule
  %% Primitives& $+,\times:\mathbb{R}\to \mathbb{R}\to \mathbb{R}$\\
  %% & $\mathcal{R}:\mathbb{R}$ (real valued parameter)\\\midrule
  %% Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
  %% \midrule
  %% Likelihood $\probability[x|p]$& $\propto\exp\left( - D\log N \right)\prod_{n\leq N}\mathcal{N}(p(i_n)|o_n)$
  %% \\\midrule
    $f_0 = $\texttt{(+ }$\mathcal{R}$\texttt{)}\\
    $f_1($\texttt{x}$) =$ \texttt{(* (* (+ x }$\mathcal{R}$\texttt{) }$\mathcal{R}$\texttt{) x)} \\
      $f_2($\texttt{x}$) =$ \texttt{(+ }$\mathcal{R}$\texttt{ ($f_1$ x))} \\
      $f_3($\texttt{x}$) = $\texttt{(* x (}$f_2$\texttt{ x))}\\
  $f_4($\texttt{x}$) = $\texttt{(}$f_0$\texttt{ (}$f_3$\texttt{ x)))}\\
  $f_5($\texttt{x}$) =$ \texttt{(* x (}$f_4$\texttt{ x))}\\
  $f_6($\texttt{x}$) = $\texttt{(}$f_0$\texttt{ (}$f_4$\texttt{ x))} (\emph{4th order polynomial})
  \\\bottomrule
\end{tabular}
\caption{Some induced DSL primitives for symbolic regression. System starts out with addition, multiplication, and real numbers, and learns functions for building 4-th order polynomials.}\label{regressionDSL}
\end{table}

\begin{figure}\centering
  \includegraphics[width = 7cm]{figures/symbolicRegression.png}
  \caption{DSL learning curve for symbolic regression.}\label{regressionLearningCurve}
\end{figure}
\subsection{String editing}\label{textSection}
